Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Proof-of-Concept VI}}

### 2014-10-04 Paper.tex
MSIZE is bytes.
Moved out old JS and Wire APIs.

We define $\boldsymbol{\sigma}_1$, the checkpoint state as the original state but with the value transferred to the recipient:
\begin{eqnarray}
\boldsymbol{\sigma}_1 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
\boldsymbol{\sigma}_1[r]_b & \equiv & \boldsymbol{\sigma}_1[r]_b + v
\end{eqnarray}

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
2G_{sstore} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
G_{sstore} & \text{otherwise}
\end{cases}$ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
\bottomrule
\end{tabular*}

@@

\section{Wire Protocol}\label{app:wire}
The wire-protocol specifies a network-level protocol for how two peers can communicate. It includes handshake procedures and the means for transferring information such as peers, blocks and transactions. Peer-to-peer communications between nodes running Ethereum clients are designed to be governed by a simple wire-protocol making use of existing Ethereum technologies and standards such as RLP wherever practical.

Ethereum nodes may connect to each other over TCP only. Peers are free to advertise and accept connections on any port(s) they wish, however, a default port on which the connection may be listened and made will be 30303.

Though TCP provides a connection-oriented medium, Ethereum nodes communicate in terms of packets. These packets are formed as a 4-byte synchronisation token (0x22400891), a 4-byte "payload size", to be interpreted as a big-endian integer and finally an N-byte \textbf{RLP-serialised} data structure, where N is the aforementioned "payload size". To be clear, the payload size specifies the number of bytes in the packet ''following'' the first 8.

There are a number of different types of message that may be sent. This ``type'' is always determined by the first entry of the structure, represented as a scalar. The structure of each message type is described below.

\begin{tabular*}{\columnwidth}[h]{rlll}
\toprule
\multicolumn{4}{c}{\textbf{00s: Session control}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
0x00 & \textsc{Hello} & & $(\text{0x}00, v \in \mathbb{P}, n \in \mathbb{P}, \mathbf{i} \in \mathbb{B}, c \in \mathbb{P}, p \in \mathbb{P}, u \in \mathbb{B}_{64})$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
This is the first packet sent over the connection, and sent once by both sides. No other messages may be sent until a \textsc{Hello} is received.
\begin{itemize}
\item $v$ is the Protocol Version. See the latest documentation for which version is current.
\item $n$ is the Network Id should be 0.
\item $\mathbf{i}$ is the Client Id and specifies the client software identity as a human-readable string (e.g. ``Ethereum(++)/1.0.0'').
\item $c$ is the client's Capabilities and specifies the capabilities of the client as a set of flags; presently three bits are used:
\begin{description}
\item[0x01] Client provides peer discovery service;
\item[0x02] Client provides transaction relaying service;
\item[0x04] Client provides block-chain querying service.
\end{description}
\item $p$ is the Listen Port and specifies the port that the client is listening on (on the interface that the present connection traverses). If 0 it indicates the client is not listening.
\item $u$ is the Unique Identity of the node and specifies a 512-bit hash that identifies this node.
\end{itemize}
}\\
\midrule
0x01 & \textsc{Disconnect} && $(\text{0x}01, r \in \mathbb{P})$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
Inform the peer that a disconnection is imminent; if received, a peer should disconnect immediately. When sending, well-behaved hosts give their peers a fighting chance (read: wait 2 seconds) to disconnect to before disconnecting themselves.
\begin{itemize}
\item $r$ is an integer specifying one of a number of reasons for disconnect:
\begin{description}
\item[0x00] Disconnect requested;
\item[0x01] TCP sub-system error;
\item[0x02] Bad protocol;
\item[0x03] Useless peer;
\item[0x04] Too many peers;
\item[0x05] Already connected;
\item[0x06] Wrong genesis block;
\item[0x07] Incompatible network protocols;
\item[0x08] Client quitting.
\end{description}
\end{itemize}
}\\
\midrule
0x02 & \textsc{Ping} & \textsc{Pong} & $(\text{0x}02)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Requests an immediate reply of \textsc{Pong} from the peer.}\\
\midrule
0x03 & \textsc{Pong} && $(\text{0x}03)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Reply to peer's \textsc{Ping} packet.}\\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlll}
\toprule
\multicolumn{4}{c}{\textbf{10s: Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
0x10 & \textsc{GetPeers} & \textsc{Peers} & $(\text{0x}10)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to enumerate some known peers for us to connect to. This should include the peer itself.}\\
\midrule
0x11 & \textsc{Peers} & & $(\text{0x}11, (a_0 \in \mathbb{B}_4, p_0 \in \mathbb{P}, i_0 \in \mathbb{B}_{64}), (a_1 \in \mathbb{B}_4, p_1 \in \mathbb{P}, i_1 \in \mathbb{B}_{64}), ...)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
Specifies a number of known peers. 
\begin{itemize}
\item $a_0$, $a_1$, ... is the node's IPv4 address, a 4-byte array that should be interpreted as the IP address $a_0[0]$.$a_0[1]$.$a_0[2]$.$a_0[3]$.
\item $p_0$, $p_1$, ... is the node's Port and is an integer.
\item $i_0$, $i_1$, ... is the node's Unique Identifier and is the 512-bit hash that serves to identify the node.
\end{itemize}
}\\
\midrule
0x12 & \textsc{Transactions} & & $(\text{0x}12, L_T(T_0), L_T(T_1), ...)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
where $L_T$ is the transaction preparation function, as specified in section \ref{ch:block}.

Specify a transaction or transactions that the peer should make sure is included on its transaction queue. The items in the list (following the first item 0x12) are transactions in the format described in the main Ethereum specification.
\begin{itemize}
\item $T_0$, $T_1$, ... are the transactions that should be assimilated.
\end{itemize}
}\\
\midrule
0x13 & \textsc{Blocks} && $(\text{0x}13, L_B(b_0), L_B(b_1), ...)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
Where $L_B$ is the block preparation function, as specified in section \ref{ch:block}.

Specify a block or blocks that the peer should know about. The items in the list (following the first item, 0x13) are blocks as described in the format described in the main specification.
\begin{itemize}
\item $b_0$, $b_1$, ... are the blocks that should be assimilated.
\end{itemize}
}\\
\midrule
0x14 & \textsc{GetChain} & \textsc{Blocks} or \textsc{NotInChain} & $(\text{0x}14, p_0 \in \mathbb{B}_{32}, p_1 \in \mathbb{B}_{32}, ..., c \in \mathbb{P})$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
Request the peer to send $c$ blocks in the current canonical block chain that are children of one of a number of given blocks, according to a preferential order with $p_0$ being the most prefered. If the designated parent is the present block chain head, an empty reply should be sent. If none of the parents are in the current canonical block chain, then a NotInChain message should be sent along with $p_n$, the least preferential parent. If no parents are passed, then a reply need not be made.
\begin{itemize}
\item $p_0$, $p_1$, ... are the SHA3 hashes of the parents of blocks that we should be informed of with a \textsc{Blocks} reply. Typically, these will be specified in increasing age (or decreasing block number).
\item $c$ is the number of children blocks of the most preferred parent that we should be informed of through the corresponding \textsc{Blocks} reply.
\end{itemize}
}\\
\midrule
0x15 & \textsc{NotInChain} && $(\text{0x}15, p \in \mathbb{B}_{32})$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Inform the peer that a particular block was not found in its block chain.
\begin{itemize}
\item $p$ is the SHA3 hash of the block that was not found in the block chain. Typically, this will be the least preferential (oldest) block hash given in a previous \textsc{GetChain} message.
\end{itemize}
}\\
\midrule
0x16 & \textsc{GetTransactions} & \textsc{Transactions} & $(\text{0x}16)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to send all transactions currently in the queue. See \textsc{Transactions}.}\\
\bottomrule
\end{tabular*}

@@

\section{Javascript API}\label{app:jsapi}

The JavaScript API provides a consistent API across multiple scenarios including each of the clients' web-based in-process \DH{}App frameworks and the out-of-process RPC-based infrastructure. All key access takes places though the special \texttt{eth} object, part of the global namespace.

\subsection{Values}
There are no special object types in the API; all values are strings. As strings, values may be of several forms, and are interpreted by the API according to a series of rules:

\begin{enumerate}
\item If the string contains only digits from 0-9, then it is interpreted as a decimal integer;
\item if the string begins with the characters \texttt{0x}, then it is interpreted as a hexadecimal integer;
\item it is interpreted as a binary string otherwise.
\end{enumerate}

The only exception to this are for parameters that expect a binary string; in this case the string is always interpreted as such.

Values are implicitly converted between integers and hashes/byte-arrays; when this happens, integers are interpreted as big-endian as is standard for Ethereum. The following forms are allowed; they are all interpreted in the same way:

\begin{enumerate}
\item \texttt{"4276803"}
\item \texttt{"0x414243"}
\item \texttt{"ABC"}
\end{enumerate}

In each case, they are interpreted as the number 4276803. The first two values may be alternated between with the additional String methods \texttt{bin()} and \texttt{unbin()}.

As byte arrays, values may be concatenated with the \texttt{+} operator as is normal for strings.

Strings also have a number of additional methods to help with conversion and alignment when switching between addresses, 256-bit integers and free-form byte-arrays for transaction data:

\begin{itemize}
\item \texttt{bin()}: Converts the string to binary format.
\item \texttt{pad(l)}: Converts the string to binary format (ready for data parameters) and pads with zeroes until it is of width \texttt{l}. Will pad to the left if the original string is numeric, or to the right if binary. If \texttt{l} is less than the width of the string, it is resized accordingly.
\item \texttt{pad(a, b)}: Converts the string to binary format (ready for data parameters) and pads with zeroes on the left size until it is of width \texttt{a}. Then pads with zeroes on the right side until it has grown to size \texttt{b}. If \texttt{b} is less hat the width of the string, it is resized accordingly.
\item \texttt{unbin()}: Converts the string from binary format to hex format.
\item \texttt{unpad()}: Converts the string from binary format to hex format, first removing any zeroes from the right side.
\item \texttt{dec()}: Converts the string to decimal format (typically from hex).
\end{itemize}

\subsection{The \texttt{eth} object}

\subsubsection{Properties}

For each such item, there is also an asynchronous method, taking a parameter of the callback function, itself taking a single parameter of the property's return value and of the same name but prefixed with get and recapitalised, e.g. \texttt{getCoinbase(fn)}.

\begin{itemize}
\item \texttt{coinbase} Returns the coinbase address of the client.
\item \texttt{isListening} Returns true if and only if the client is actively listening for network connections.
\item \texttt{isMining} Returns true if and only if the client is actively mining new blocks.
\item \texttt{gasPrice} Returns the client's present price of gas.
\item \texttt{key} Returns the special key-pair object corresponding to the preferred account owned by the client.
\item \texttt{keys} Returns a list of the special key-pair objects corresponding to all accounts owned by the client.
\item \texttt{peerCount} Returns the number of peers currently connected to the client.
\end{itemize}

\subsubsection{Synchronous Getters}
For each such item, there is also an asynchronous method, taking an additional parameter of the callback function, itself taking a single parameter of the synchronous method's return value and of the same name but prefixed with get and recapitalised, e.g. \texttt{getBalanceAt(a, fn)}.

\begin{itemize}
\item \texttt{balanceAt(a)} Returns the balance of the account of address given by the address \texttt{a}.
\item \texttt{storageAt(a, x)} Returns the value in storage at position given by the number x of the account of address given by the address \texttt{a}.
\item \texttt{txCountAt(a)} Returns the number of transactions send from the account of address given by \texttt{a}.
\item \texttt{isContractAt(a)} Returns true if the account of address given by \texttt{a} has associated code.
\end{itemize}

\subsubsection{Transactions}

\begin{itemize}
\item \texttt{create(sec, xEndowment, bCode, xGas, xGasPrice, fn)} Creates a new contract-creation transaction, given parameters:
\begin{itemize}
\item \texttt{sec}, the secret-key for the sender;
\item \texttt{xEndowment}, the number equal to the account's endowment;
\item \texttt{bCode}, the binary string (byte array) of EVM-bytecode for the initialisation of the account;
\item \texttt{xGas}, the number equal to the amount of gas to purchase for the transaction (unused gas is refunded);
\item \texttt{xGasPrice}, the number equal to the price of gas for this transaction. Returns the special address object representing the new account; and
\item \texttt{fn}, the callback function, called on completion of the transaction.
\end{itemize}
\item \texttt{transact(sec, xValue, aDest, bData, xGas, xGasPrice, fn)} Creates a new message-call transaction, given parameters:
\begin{itemize}
\item \texttt{sec}, the secret-key for the sender;
\item \texttt{xValue}, the value transferred for the transaction (in Wei);
\item \texttt{aDest}, the address representing the destination address of the message;
\item \texttt{bData}, the binary string (byte array), containing the associated data of the message;
\item \texttt{xGas}, the amount of gas to purchase for the transaction (unused gas is refunded);
\item \texttt{xGasPrice}, the price of gas for this transaction; and
\item \texttt{fn}, the callback function, called on completion of the transaction.
\end{itemize}
\end{itemize}

\subsubsection{Events}

\begin{itemize}
\item \texttt{watch(a, fn)}: Registers \texttt{fn} as a callback for whenever anything about the state of the account at address \texttt{a} changes, and also on the initial load.
\item \texttt{watch(a, x, fn)}: Registers \texttt{fn} as a callback for whenever the storage location \texttt{x} of the account at address \texttt{a} changes, and also on the initial load.
\item \texttt{newBlock(fn)}: Registers \texttt{fn} as a callback for whenever the state changes, and also on the initial load.
\end{itemize}

\subsubsection{Misc}

\begin{itemize}
\item \texttt{secretToAddress(a)}: Determines the address from the secret key \texttt{a}.
\end{itemize}

>>>>

We define $\boldsymbol{\sigma}_1$, the checkpoint state as the original state but with the value transferred to the recipient:
\begin{eqnarray}
\boldsymbol{\sigma}_1 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
\boldsymbol{\sigma}_1[r]_b & \equiv & \boldsymbol{\sigma}[r]_b + v
\end{eqnarray}

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
2G_{sstore} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
G_{sstore} & \text{otherwise}
\end{cases}$ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
\bottomrule
\end{tabular*}

@@

@@
